---
title: "Final Project_Melissa Howlett"
date: "March 8, 2017"
---

STEP 1: get data and confirm it's in the right format

```{r}
#get data
faostat=read.csv("https://github.com/howlem/projectData/raw/master/Production_Crops_E_All_Data.csv",stringsAsFactors = F)

```

```{r}
#check that it is a data frame
is.data.frame(faostat)
```

STEP 2: install the libraries needed later

```{r}
#install libraries
library(DescTools)
library(ggplot2)
library(moments)
library(stringr)
library(scales)
library(car)
library(utils)
library(rgdal)
library(RColorBrewer)
library(classInt)
library(cluster)

```

STEP 3: see what data we have

```{r}
#review data called
head(faostat)
str(faostat)
names(faostat)
```

STEP 4: isolate each variable and aggregate crop data by country

```{r}
##YIELD##
#just keep yield data
yield_1=faostat[faostat$'Element'!='Production',]
yield=yield_1[yield_1$'Element'!='Area harvested',]
head(yield)

#aggregate by country
yield=aggregate(yield$'Y2016',by=list(yield$'Area'),sum)
names(yield)=c('country','TotYield2016')
yield

str(yield)
```

```{r}
##PRODUCTION##
#just keep production data
prod_1=faostat[faostat$'Element'!='Yield',]
prod=prod_1[prod_1$'Element'!='Area harvested',]
head(prod)

#aggregate by country
prod=aggregate(prod$'Y2016',by=list(prod$'Area'),sum)
names(prod)=c('country','TotProd2016')
prod
```

```{r}
##AREA HARVESTED##
#just keep area harvested data
area_harv_1=faostat[faostat$'Element'!='Yield',]
area_harv=area_harv_1[area_harv_1$'Element'!='Production',]
head(area_harv)

#aggregate by country
area_harv=aggregate(area_harv$'Y2016',by=list(area_harv$'Area'),sum)
names(area_harv)=c('country','TotAreaHarv2016')
area_harv
```

STEP 5: conduct univariate analysis of yield

```{r}
#check data types
str(yield)
```

```{r}
#centrality
summary(yield$'TotYield2016')
```

```{r}
#dispersion
sd(yield$'TotYield2016',na.rm=T)
```

```{r}
#coefficient of variation
sd(yield$'TotYield2016',na.rm=T)/mean(yield$'TotYield2016',na.rm=T)
```

```{r}
#skewness
skewness(yield$'TotYield2016',na.rm=T)
```

```{r}
#kurtosis
kurtosis(yield$'TotYield2016',na.rm=T)
Kurt(yield$'TotYield2016',na.rm=T)
```

```{r}
##histogram
#keep non missing
data=yield[is.finite(yield$'TotYield2016'),]

#making numeric
as.numeric(data$'TotYield2016')

#histogram
hist(data$'TotYield2016')
```

```{r}
#histogram of log
hist(log(data$'TotYield2016'))
```

```{r}
#saving mean and sd
mnVar=mean(data$TotYield2016,na.rm=T)
sdVar=sd(data$TotYield2016,na.rm=T)
```

```{r}
#outliers
summary(yield$'TotYield2016')
```

```{r}
#boxplot
(bp=boxplot(yield$'TotYield2016',horizontal=T))
```

```{r}
#boxplot of log
(bp_log=boxplot(log(yield$'TotYield2016')))
```

```{r}
#show values of both boxplots
bp$stats
bp_log$stats
```

```{r}
#show number of outliers
length(bp$out)
length(bp_log$out)
```

STEP 6: conduct bivariate analysis of yield, using production and area harvested

```{r}
#merge yield, production, and area harvested data into one dataset
combinedYP=merge(yield,prod)
combinedYPAH=merge(combinedYP,area_harv)
head(combinedYPAH)
```

```{r}
#correlation
res<-cor(combinedYPAH[,c(2:4)],use='complete.obs')
round(res,2)
```

```{r}
#correlation of log
res_log<-cor(log(combinedYPAH[,c(2:4)]),use='complete.obs')
round(res_log,2)
```

```{r}
#scatterplot
scatterplotMatrix(combinedYPAH[,c(2:4)],
                  main="Simple Scatterplot Matrix")
```

```{r}
#scatterplot of log
scatterplotMatrix(log(combinedYPAH[,c(2:4)]),
                  main="Simple Scatterplot Matrix")
```

```{r}
#non-parametric correlation
cor(combinedYPAH[,c(2:4)],use='complete.obs',method='spearman')
```

STEP 7: Bring in map data

```{r}
##MAP DATA##
#get zip file from github
compressedMap="https://github.com/howlem/projectData/blob/master/worldMap/worldMap.shp?raw=true"
```

```{r}
#unzip the file
temp=tempfile()
download.file(compressedMap,temp)
unzip(temp)
```

```{r}
#select the map (shp file) needed
thefile=file.path('worldMap','worldMap.shp')
worldMap <- rgdal::readOGR(thefile,stringsAsFactors=F) # use the names
```

```{r}
#check it out
head(worldMap)
```
STEP 8: Limit map to Africa

```{r}
#only keep African countries
worldMap=worldMap[worldMap$REGION==2,]
```

STEP 9: Combine shp file with FAO data

```{r}
#see what variables are present
names(worldMap)

```

```{r}
#rename "NAME" column to "country"" so can merge
names(worldMap)[names(worldMap)=='NAME'] <- 'country'
```

```{r}
#merge the two datasets into one, by country
YPAHforMap <- merge(worldMap,combinedYPAH,
                    by=c("country"))
```

```{r}
#see what happened
head(YPAHforMap)
```

STEP 10: Map yield data

```{r}
#identify variable to plot
varToPlot=YPAHforMap$TotYield2016

#plot yield data on map of Africa
numberOfClasses = 8
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPlot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)

legendText="Total Yield of all crops in 2016"
shrinkLegend=0.5
title="Total 2016 Yield by Country in Africa"

plot(worldMap,col='gray',main=title)
plot(YPAHforMap, col=colorPallette, border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

STEP 11: Map production data

```{r}
#identify variable to plot 
varToPlot=YPAHforMap$TotProd2016

#plot production data on map of Africa
numberOfClasses = 8
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPlot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)

legendText="Total Production of all crops in 2016"
shrinkLegend=0.5
title="Total 2016 Production by Country in Africa"

plot(worldMap,col='gray',main=title)
plot(YPAHforMap, col=colorPallette, border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

STEP 12: Map area harvested data

```{r}
#identify variable to plot 
varToPlot=YPAHforMap$TotAreaHarv2016

#plot area harvested data on map of Africa
numberOfClasses = 8
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPlot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)

legendText="Total Area Harvested of all crops in 2016"
shrinkLegend=0.5
title="Total 2016 Area Harvested by Country in Africa"

plot(worldMap,col='gray',main=title)
plot(YPAHforMap, col=colorPallette, border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```