---
title: "Final Project_Melissa Howlett"
date: "February 1, 2017"
---

STEP 1: get data and confirm it's in the right format

```{r}

#get data
faostat=read.csv("https://github.com/howlem/projectData/raw/master/Production_Crops_E_All_Data.csv",stringsAsFactors = F)

#check is data frame
is.data.frame(faostat)

```

STEP 2: install the libraries needed later

```{r}
#install libraries
library(DescTools)
library(ggplot2)
library(moments)
library(stringr)
library(scales)

```

STEP 3: see what data we have

```{r}
#review data called
head(faostat)
str(faostat)
names(faostat)
```

STEP 4: isolate each variable and aggregate crop data by country

```{r}
##YIELD##
#just keep yield data
yield_1=faostat[faostat$'Element'!='Production',]
yield=yield_1[yield_1$'Element'!='Area harvested',]
head(yield)

#aggregate by country
yield=aggregate(yield$'Y2016',by=list(yield$'Area'),sum)
names(yield)=c('country','TotYield2016')
yield

str(yield)

##PRODUCTION##
#just keep production data
prod_1=faostat[faostat$'Element'!='Yield',]
prod=prod_1[prod_1$'Element'!='Area harvested',]
head(prod)

#aggregate by country
prod=aggregate(prod$'Y2016',by=list(prod$'Area'),sum)
names(prod)=c('country','TotProd2016')
prod

##AREA HARVESTED##
#just keep area harvested data
area_harv_1=faostat[faostat$'Element'!='Yield',]
area_harv=area_harv_1[area_harv_1$'Element'!='Production',]
head(area_harv)

#aggregate by country
area_harv=aggregate(area_harv$'Y2016',by=list(area_harv$'Area'),sum)
names(area_harv)=c('country','TotAreaHarv2016')
area_harv
```

STEP 5: conduct univariate analysis of yield

```{r}

#check data types
str(yield)

#centrality
summary(yield$'TotYield2016')

#dispersion
sd(yield$'TotYield2016',na.rm=T)

##coefficient of variation
sd(yield$'TotYield2016',na.rm=T)/mean(yield$'TotYield2016',na.rm=T)

#skewness
skewness(yield$'TotYield2016',na.rm=T)

#kurtosis
kurtosis(yield$'TotYield2016',na.rm=T)
Kurt(yield$'TotYield2016',na.rm=T)

#histogram
##keep non missing
data=yield[is.finite(yield$'TotYield2016'),]

##making numeric
as.numeric(data$'TotYield2016')

##histogram
hist(data$'TotYield2016')

##histogram of log
hist(log(data$'TotYield2016'))

#saving mean and sd
mnVar=mean(var,na.rm=T)
sdVar=sd(var,na.rm=T)

#outliers
summary(yield$'TotYield2016')

##boxplot
(bp=boxplot(yield$'TotYield2016',horizontal=T))

##boxplot of log
(bp_log=boxplot(log(yield$'TotYield2016')))

#show values of both boxplots
bp$stats
bp_log$stats

#show number of outliers
length(bp$out)
length(bp_log$out)

```

STEP 6: conduct bivariate analysis of yield, using production and area harvested

```{r}

#merge yield, production, and area harvested data into one dataset
combinedYP=merge(yield,prod)
combinedYPAH=merge(combined,area_harv)
head(combinedYPAH)

#correlation
res<-cor(combinedYPAH[,c(2:4)],use='complete.obs')
round(res,2)

#correlation of log
res_log<-cor(log(combinedYPAH[,c(2:4)]),use='complete.obs')
round(res_log,2)

#scatterplot
library(car)
scatterplotMatrix(combinedYPAH[,c(2:4)],
                  main="Simple Scatterplot Matrix")

#scatterplot of log
scatterplotMatrix(log(combinedYPAH[,c(2:4)]),
                  main="Simple Scatterplot Matrix")

#non-parametric correlation
cor(combinedYPAH[,c(2:4)],use='complete.obs',method='spearman')

```

```{r}
##MAP DATA
#get zip file from github
compressedMap="https://github.com/howlem/projectData/raw/master/worldMap%20(1).zip"

#unzip the file
library(utils)
temp=tempfile()
download.file(compressedMap,temp)
unzip(temp)

```

```{r}
#select the map (shp file) needed
library(rgdal)
thefile=file.path('worldMap','worldMap.shp')
worldMap <- rgdal::readOGR(thefile,stringsAsFactors=F) # use the names

#check it out
head(worldMap)

```

```{r}


```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
