---
title: "Final Project_Melissa Howlett"
date: "February 1, 2017"
---

STEP 1: get data and confirm it's in the right format

```{r}

#get data
faostat=read.csv("https://github.com/howlem/projectData/raw/master/Production_Crops_E_All_Data.csv",stringsAsFactors = F)

#check is data frame
is.data.frame(faostat)

```

STEP 2: install the libraries needed later

```{r}
#install libraries
library(DescTools)
library(ggplot2)
library(moments)
library(stringr)
library(scales)

```

STEP 3: see what data we have

```{r}
#review data called
head(faostat)
str(faostat)
names(faostat)
```

STEP 4: isolate each variable and aggregate crop data by country

```{r}
##YIELD##
#just keep yield data
yield_1=faostat[faostat$'Element'!='Production',]
yield=yield_1[yield_1$'Element'!='Area harvested',]
head(yield)

#aggregate by country
yield=aggregate(yield$'Y2016',by=list(yield$'Area'),sum)
names(yield)=c('country','TotYield2016')
yield

str(yield)

##PRODUCTION##
#just keep production data
prod_1=faostat[faostat$'Element'!='Yield',]
prod=prod_1[prod_1$'Element'!='Area harvested',]
head(prod)

#aggregate by country
prod=aggregate(prod$'Y2016',by=list(prod$'Area'),sum)
names(prod)=c('country','TotProd2016')
prod

##AREA HARVESTED##
#just keep area harvested data
area_harv_1=faostat[faostat$'Element'!='Yield',]
area_harv=area_harv_1[area_harv_1$'Element'!='Production',]
head(area_harv)

#aggregate by country
area_harv=aggregate(area_harv$'Y2016',by=list(area_harv$'Area'),sum)
names(area_harv)=c('country','TotAreaHarv2016')
area_harv
```

STEP 5: conduct univariate analysis of yield

```{r}

#check data types
str(yield)

#centrality
summary(yield$'TotYield2016')

#dispersion
sd(yield$'TotYield2016',na.rm=T)

##coefficient of variation
sd(yield$'TotYield2016',na.rm=T)/mean(yield$'TotYield2016',na.rm=T)

#skewness
skewness(yield$'TotYield2016',na.rm=T)

#kurtosis
kurtosis(yield$'TotYield2016',na.rm=T)
Kurt(yield$'TotYield2016',na.rm=T)

#histogram
##keep non missing
data=yield[is.finite(yield$'TotYield2016'),]

##making numeric
as.numeric(data$'TotYield2016')

##histogram
hist(data$'TotYield2016')

##histogram of log
hist(log(data$'TotYield2016'))

#saving mean and sd
mnVar=mean(data$TotYield2016,na.rm=T)
sdVar=sd(data$TotYield2016,na.rm=T)

#outliers
summary(yield$'TotYield2016')

##boxplot
(bp=boxplot(yield$'TotYield2016',horizontal=T))

##boxplot of log
(bp_log=boxplot(log(yield$'TotYield2016')))

#show values of both boxplots
bp$stats
bp_log$stats

#show number of outliers
length(bp$out)
length(bp_log$out)

```

STEP 6: conduct bivariate analysis of yield, using production and area harvested

```{r}

#merge yield, production, and area harvested data into one dataset
combinedYP=merge(yield,prod)
combinedYPAH=merge(combinedYP,area_harv)
head(combinedYPAH)

#correlation
res<-cor(combinedYPAH[,c(2:4)],use='complete.obs')
round(res,2)

#correlation of log
res_log<-cor(log(combinedYPAH[,c(2:4)]),use='complete.obs')
round(res_log,2)

#scatterplot
library(car)
scatterplotMatrix(combinedYPAH[,c(2:4)],
                  main="Simple Scatterplot Matrix")

#scatterplot of log
scatterplotMatrix(log(combinedYPAH[,c(2:4)]),
                  main="Simple Scatterplot Matrix")

#non-parametric correlation
cor(combinedYPAH[,c(2:4)],use='complete.obs',method='spearman')

```

STEP 7: Bring in map data

```{r}
##MAP DATA##
#get zip file from github
compressedMap="https://github.com/howlem/projectData/raw/master/worldMap%20(1).zip"

#unzip the file
library(utils)
temp=tempfile()
download.file(compressedMap,temp)
unzip(temp)

```

```{r}
#select the map (shp file) needed
library(rgdal)
thefile=file.path('worldMap','worldMap.shp')
worldMap <- rgdal::readOGR(thefile,stringsAsFactors=F) # use the names

#check it out
head(worldMap)

```

```{r}
#only keep African countries
worldMap=worldMap[worldMap$REGION==2,]
```

STEP 8: Combine it with FAO data

```{r}
#see what variables are present
names(worldMap)

#rename "NAME" column to "country"" so can merge
names(worldMap)[names(worldMap)=='NAME'] <- 'country'

#merge the two datasets into one, by country
YPAHforMap <- merge(worldMap,combinedYPAH,
                    by=c("country"))

#see what happened
head(YPAHforMap)

```

```{r}
names(YPAHforMap)

```

```{r}

library(RColorBrewer)
library(classInt)

varToPlot=YPAHforMap$TotYield2016

numberOfClasses = 8
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPlot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)

legendText="Total Yield of all crops in 2016"
shrinkLegend=0.5
title="Total 2016 Yield by Country in Africa"

plot(worldMap,col='gray',main=title)
plot(YPAHforMap, col=colorPallette, border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

```{r}
names(YPAHforMap)
```


```{r}
test=YPAHforMap@data[,c(1,12:14)]
```


```{r}

combinedYPAH=combinedYPAH[complete.cases(combinedYPAH),]

library(cluster)

YPAHsimi=daisy(combinedYPAH[,c(2:4)],metric="euclidean",
                   stand=TRUE)

YPAHcluster=hclust(YPAHsimi,method = 'average')

plot(YPAHcluster,cex=0.3)
rect.hclust(YPAHcluster, k = 3,border = c('orange','blue','red'))

```

```{r}
combinedYPAH[,c(2:4)]
```

```{r}
#numerical clustering
YPAHsil3=silhouette(cutree(YPAHcluster,k=3),YPAHsimi)
clust3solution=data.frame(matrix(YPAHsil3,ncol = 3,dimnames =dimnames(YPAHsil3)))
row.names(clust3solution)=combinedYPAH$country
```

```{r}
clust3solution
```
```{r}
#identify which items are difficult to cluster
(bad3=clust3solution[clust3solution$sil_width<0,])
```

```{r}

```

```{r}

#save solution
solution=clust3solution

#create country column
solution$country=row.names(clust3solution)
row.names(solution)=NULL #reset row names

#merge datasets
combinedYPAH_updated=merge(YPAHforMap,solution)

head(combinedYPAH_updated)

```

```{r}

varToPlot=combinedYPAH_updated$cluster

numberOfClasses = length(levels(varToPlot)) 
colorForScale='Set2'
colors = brewer.pal(numberOfClasses, colorForScale)

legendText="Clustered - Yield, Production, and Area Harvested"
shrinkLegend=0.4
title="Yield, Production, and Area Harvested in Africa (2016)"

plot(worldMap,col='black',main=title)
plot(combinedYPAH_updated, 
     col = colors[varToPlot],border='grey',add=T)

legend('left', legend = levels(varToPlot), 
       fill = colors, 
       cex = shrinkLegend, 
       bty = "n",
       title=legendText)

```








```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
